Kinematics \& Chemistry of Halo Substructures: The Vicinity of the Virgo Over-Density

### 2012-05-28 2011-sgr/AJ-version/referee-response-1.txt

Firstly, the authors wish to thank the referee for their constructive comments. Apologies for the delay in responding to feedback.

Introduction:

1) Some references to the substructures alluded to in the first sentence should be provided.

Added reference to the Helmi 2008 review paper (2008A&ARv..15..145H), and references therein.

2) The term "admits" would more appropriately be replaced with "emphasizes".

Corrected in the revised manuscript.

Section 2. Target Selection

1) Some explicit statement should be included to the effect that, while the selection box favors a metal poor population at 40 kpc, it will also contain (and not be significantly biased against) similar metal poor stars at half that distance (e.g. the VSS).

Added "This selection box favors a metal-poor population at Sgr stream-like distances ($\sim40$ kpc) but will also contain similarly metal-poor stars at VOD/VSS distances ($\sim20$ kpc) without being significantly biased."

2) Some reference should be made to Table 1, and reasons given for choosing the particular fields observed. Are these fields part of a larger survey, or were they optimized for the VOD/VSS?

Edited to include this reference, and noted that fields were chosen to be within the VOD/Sgr region.

Section 3. Observations

1) The spectral resolution in the red arm (where the Ca lines of interest are) should be provided.

Added (R = 4400) in the revised manuscript.

2) Note that the reddening maps of Schlegel et al. (1998) have been significantly recalibrated by Schafly et al. (2010, 2011). These corrections (~14%) should be taken into account.

Corrections made using Schlafly (2010, 2011) corrections and assuming a Fitzpatrick (1999) profile where $R_V$ = 3.1 - these changes are also reflected in the manuscript.

Section 4. Dwarf/Giant separation

1) The colors in Figure 2 are somewhat difficult to sort out. Colors distinguishable by 99% of the population include black, white, and the primary colors (blue, red, and yellow) and, when it is necessary to resort to colors,  these should always be used first.

Primary colors have been used to identify metallicity, and the size of these points has been increased for clarity.

Section 6. Metallicities

1) "ascertained" is better replaced with "measured".

Corrected in the revised manuscript.

Section 7.1 Substructure Identification

1) The Gaussian halo distribution in Figure 3 doesn't look properly normalized. I.e. the area under the curve appears to be less than that under the histogram of observational velocities.  If corrected, then features B and C would then appear to be considerably less significant, though the simulation results will be unchanged.

The referee is correct in that the area under the curve is less than the area under the histogram of observational velocities. This is because the Gaussian distribution is meant to represent the underlying halo population present. Originally we have started with a Gaussian sample of 178 giants and we have removed significant (> 2.5 sigma) kinematic deviations (substructure) from the population, resulting in the Gaussian shown. This is noted in the figure caption, however it has been made more explicit;

"The Gaussian represents the halo contribution; it is normalised such that the integral equals the number of observed stars excluding those outside a 2.5-$\sigma$ excess."

2) The term "bandwidth" for the width or sigma of a normal distribution is not really appropriate. This should simply be called "width" or "sigma", both here and in Section 7.2.1

Corrected in Sections 7.1 and 7.2.1.

3) The simulation plot (Figure 4) is interesting, but somewhat difficult to interpret. For one thing, it is difficult to see individual data points superposed on the differently colored grid blocks. This figure should include a separate panel showing just the data points by themselves. Having done that, it is probably not necessary to show them superimposed on the simulations. Here again color is an issue. From context, I infer that the black (or very dark) blocks show regions in [Fe/H] vs V_GSR where the simulations rarely or never produced any stars. But this isn't clear from the color-bar, both ends of which are very dark. Moreover, the red through white through blue scaling is, to my eye, somewhat non-monotonic. In this instance, a grey scale plot would be much more readily interpreted, with very light indicating numerous simulation counts, and very dark indicating infrequent occurrences.

This plot has been separated into two panels as described, and is now in greyscale (and is noticeably easier to interpret).

Section 7.2.1

1) B_odot and Lambda_odot should be defined here (in words), rather than forcing the reader to consult LM10.

Added; "The $10^5$ Sgr debris particles are represented along the best-fit great-circle ($\Lambda_\odot$, $B_\odot$) where the longitudinal coordinate $\Lambda_\odot$ is zero in the direction of the Sgr core and increases along the trailing debris."

2) The scaling of the LM10 models should be described.  Should the particles correspond to the Sgr stream star for star? Or does each particle represent some number of stars, presumably scattered around each particle's position?  The upshot is that, while one can make arguments about relative likelihood, I don't think one can make definitive statements based on the fact that only 2 particles (stars) are found in the observed region using a prolate model. If each particle represents hundreds or thousands of stars, then Sgr stars could still outnumber all other halo stars in this region, and the population found here might well be consistent with the model. It might be interesting to examine the distribution of Sgr giants among the four observed fields. Are they distributed more or less uniformly, or is there a significant gradient? If not, this may be an additional reason to discount the prolate model, though not a conclusive one.

This has been addressed in the revised manuscript; "These $10^5$ particles are proportional to the stream density, implying a negligible stream concentration in this observable region. When the number of field configurations is accounted for, our observed Sgr K-giants are uniformly distributed across the fields."

3) Once again, Figure 6 is a bit difficult to interpret due to the colors used. Black, blue, red, and yellow would be much easier for virtually all readers.

Altered.

4) "forecast" is better replaced with "predict".

Corrected in the revised manuscript.

5) I am unclear as to why the K giants with highly negative velocities in the spherical and oblate models are "early". This should be explained or deleted.

The term was implying earlier (younger) K-giants, however the sentence did not contribute significant weight to the reasoning for why extremely negative velocity stars are not found. The sentence has been deleted.

6) I am puzzled by the "kinematic disruption may result from bifurcation". Should this have been the other way around? And in the following sentence, I don't think the authors can reasonably say that an unmodeled disruption scenario would yield "precisely" the broadening they see in their data.

Corrected, and removed the assumption about the unmodeled disruption scenario.

Section 7.2.2

1) The metallicity histogram in Figure 8 shows all stars with V_GSR < 0. However, in the text the authors refine their mean metallicity estimate using "only stars attributed to Feature A". How is this attribution made? By velocity? What limits? Is it the gray region in Figure 3? The limits should be explicitly indicated.

The region has been explicitly stated; "... the stars attributed to Feature A (as defined by the shaded region in Figure \ref{fig:velocity-histogram}, between $-140 < V_{gsr} < -30$ km s$^{-1}$) this ..."

2) One thing that isn't clear from Figure 4 (for the reasons given above) is what happens if one plots a velocity histogram for the stars in Figure 8 with [Fe/H] < -1.8.  Do they still show a bias towards highly negative velocities, or could they largely be halo contaminants?

The velocity histogram of our Figure 8 members with [Fe/H] < -1.8 peaks at $V_{gsr}$ = -115 km/s, and are generally biased towards highly negative velocities. This has been addressed in the revised manuscript: "Although there may be some halo contamination in our sample, in general our more metal-poor Feature A members have higher negative velocities, indicative of Sgr stream membership."

3) I think the last paragraph of this section is too understated. The target selection criteria certainly would have excluded the highest metallicity members of the stream. At a minimum, "slightly biased" should be "somewhat biased".

Corrected in the revised manuscript. It is also acknowledged in Sections 2 and 9 that our selection methodology is likely biased.

4) Section 7.3 - "calculated" should be "measured".

Corrected in the revised manuscript.

5) Concerning the spread in metallicity in the VSS indicating a dwarf galaxy rather than an isolated cluster, this is no longer such a convincing argument. We now know of several globular clusters with multiple populations.

Mentioned in the revised manuscript; "This led \citet{Duffau;et-al_2006} to infer the progenitor of the VSS was likely a dSph galaxy, as a large dispersion in $\mbox{[m/H]}$ indicates self-enrichment and implies the structure is a galaxy and not an isolated stellar cluster \--- although some globular clusters are also now known to contain sizeable dispersions in $\mbox{[m/H]}$."

6) The last sentence in the first paragraph should be expanded. How many stars are identified with the VSS, and what fraction of all stars with similar velocities does this represent?

Corrected in the revised manuscript.

Section 8

1) Some mention should be made in the abstract of the discovery of the carbon stars, if only to enable other researchers to search for and find this work.

Added 'carbon' to the star keywords and added "We also present data on five carbon stars which were discovered in our sample." as the final sentence in the abstract.

Section 9

1) The statement "a prolate dark halo predicts no debris" should be amended to something like "relatively little debris". Similarly, unless some argument can be made concerning the scaling of the models or the distribution of Sgr stars among the observed fields, it is presumptuous to conclude that "we would not expect Sgr debris to dominate our sample".

Amended:
"A prolate dark halo predicts relatively little debris in our observed fields. If the spheroid is prolate and LJM05 presents an accurate representation, Sgr debris may not dominate our sample (as it does so) even if we were $\Delta{}B_\odot\sim10^\circ$ closer to the great circle best-fit."

2) I think the last half of the last sentence is true, but perhaps not the first half. As noted above, looking at the velocity distribution of the most metal poor stars is required before one can conclude that they're not just halo contaminants.

Addressed in Section 7.2.2 of the revised manuscript and edited the final sentence: "These observations also demonstrate that the Sagittarius stream likely hosts a substantially larger abundance range than previously undiscovered, including a previously uncovered metal-poor population of K-giants."

### 2012-05-28 2011-sgr/AJ-version/sgr_vod_revised_2011.tex
Initial commit.

\author{Andrew R. Casey, Stefan Keller and Gary Da Costa}
\affil{Research School of Astronomy \& Astrophysics,
	Australian National University, Mt Stromlo Observatory, via Cotter Rd, Weston, ACT 2611, Australia}

\begin{abstract}
We present observations obtained with the AAT's 2dF wide field spectrograph AAOmega of K-type stars located within a region of the sky which contains the Virgo Over-Density and the leading arm of the Sagittarius Stream.  On the basis of the resulting velocity histogram we isolate populations in these overlapping regions, including Sagittarius and previously discovered Virgo groups. Through comparisons with \textit{N}-body models of the Galaxy-Sagittarius interaction, we find a tri-axial dark matter halo is favoured and we exclude a prolate shape. This result is contradictory with other observations along the Sagittarius leading arm, which typically favour prolate models.  We have also uncovered K-giant members of Sagittarius that are notably more metal poor ($\langle[$Fe/H$]\rangle = -1.7\,\pm\,0.3$) than previous studies. This suggests a significantly wider metallicity distribution exists in the Sagittarius Stream than formerly considered. We also present data on five serendipitous carbon stars which were uncovered in our sample.
\end{abstract}

\keywords{Galaxy: halo, structure --- Individual: Sagittarius, Virgo Stellar Stream --- Stars: K-giants}

@@

The proportion of substructure recently uncovered in the Galaxy has highlighted the crucial involvement accretion has played in the formation of the Milky Way \citep{Bell;et-al_2008}. Properties of these stellar structures allow us to probe the formation mechanisms and history of the Galaxy. Recent studies \citep{Carollo;et-al_2007, Carollo;et-al_2010} have suggested multiple evolution methods are required for galaxy formation to reconcile observational evidence, although this is a subject of ongoing debate \citep{Schoenrich;et-al_2010}. Regardless, the dissipation-less merging paradigm is widely accepted, and consistent with favoured Cold Dark Matter ($\Lambda$CDM) cosmology models. Through the examination of ongoing accretion events in the Milky Way and fossils from previous mergers, we can trace the evolution of the outer most regions of the Galaxy \citep[e.g.,][]{Helmi;White_2001}.

@@

Unquestionably the most prominent ongoing accretion event within the Milky Way is that of the Sagittarius (Sgr) dwarf Spheroidal (dSph) galaxy. Originally uncovered by \citet{Ibata;et-al_1994} as a co-moving group of K- and M-type giants, the tidal tails of Sgr circle our Galaxy. As such they have been extensively traced with red-clump stars \citep{Majewski;et-al_1999}, carbon stars \citep{Totten;Irwin_1998, Ibata;et-al_2001}, RR Lyrae stars \citep{Ivezic;et-al_2000, Vivas;et-al_2005, Keller;et-al_2008, Watkins;et-al_2009, Prior;et-al_2009b}, A-type stars \citep{Newberg;et-al_2003}, BHB stars \citep{Ruhland;et-al_2011} and K/M-giants \citep{Majewski;et-al_2003,  Yanny;et-al_2009, Keller;Yong;Da_Costa_2010}. With spatial and kinematical information, tracers originating from the host system can be unequivocally identified. This is because they remain dynamically cold, and are identifiable as kinematic substructures long after they are stripped from their progenitor \citep[for example][]{Ibata;Lewis_1998, Helmi;White_1999}.

Stellar tracers within these tails are kinematically sensitive to the galactic potential. This has led various groups to model the Sgr interaction with different dark matter profiles. \citet{Martinez-Delgado;et-al_2004} traced the Northern leading arm and found a near spherical or oblate ($q \approx 0.85$) dark matter halo best represented the observed debris, which coincided with the findings of \citet{Ibata;et-al_2001}. In contrast, \citet{Helmi_2004} found evidence in the Sgr leading debris that most favoured a prolate ($q = 1.25$) halo. \citet{Vivas;et-al_2005} found that either a prolate or spherical model of  \citet{Helmi_2004} would fit their observed RR Lyrae data, rather than those of an oblate model. \citet{Johnston;et-al_2005} later pointed out that no prolate model can reproduce the orbital pole precession of the Sgr debris, but an oblate potential could. \citet[hereafter LJM05]{Law;et-al_2005} performed simulations using data of the Sgr debris from the 2-Micron All Sky Survey (2MASS) catalogue and found that the kinematics of leading debris was best fit by prolate halos, whereas the trailing debris typically favoured oblate halos.

\citet{Belokurov;et-al_2006} found an apparent bifurcation within the Sgr debris, which \citet{Fellhauer;et-al_2006} argued can only result from a dark halo having a near spherical shape. \citet{Law;et-al_2009} introduced a tri-axial model with a varying flattening profile $q$, which replicates the orbital precession seen and matched kinematic observations of the Sgr debris. However \citet{Law;Majewski_2010} (hereafter LM10) conceded this may be a purely numerical solution as tri-axial halos are dynamically unstable, and emphasizes more kinematic measurements in other regions of the Sgr stream are required.

After the Sgr stellar stream, the Virgo Over-Density (VOD) is arguably the next most significant substructure within our Galaxy. The first over-density in the vicinity of the VOD was observed as a group of RR Lyrae stars by the QUEST survey \citep{Vivas;et-al_2001}. The collaboration later named this the ``$12\fh4$ clump'' \citep{Zinn;et-al_2004}. The broad nature of the VOD was later uncovered from the SDSS catalogue as a diffuse over-density of main-sequence turnoff stars centered at $r_\odot \sim18$ kpc \citep[which ][dubbed as S297+63-20.5]{Newberg;et-al_2002}. The nomenclature on the substructure names within this region is varied, however when referring to the VOD we are discussing the spatial over-density of stars within the region, separate from any co-moving groups.

The difficulty arises in accurately distinguishing the VOD. There are multiple substructures along this line of sight. \citet{Duffau;et-al_2006} took observations of BHB and RR Lyrae within the ``$12\fh4$ clump'' and found a common velocity of $V_{GSR} = 99.8$ km s$^{-1}$ with $\sigma_{v} = 17.3$ km s$^{-1}$. It should be noted that the dispersion in kinematics measured by \citet{Duffau;et-al_2006} is essentially their velocity precision, so the substructure may possess a much smaller $\sigma_v$. This co-moving group was coined the Virgo Stellar Stream (VSS) and thus differentiated it from the broad spatial over-density. This distinction from the VOD was somewhat strengthened with new distance measurements which placed the VOD centroid at $r_\odot = 16$ kpc \citep{Juric;et-al_2008, Keller_2010}, and the VSS 3 kpc further away \citep{Duffau;et-al_2006}. Although when considered in the light of systematic and observational uncertainties, this is of marginal significance. Moreover \citet{Juric;et-al_2008} suggests the VOD may extend between $r_\odot = 6$ to 20 kpc, which further complicates the matter of distance separation.

The relationship between the VSS and the S297+63-20.5 over-density is still unclear. \citet{Newberg;et-al_2007} found a kinematic signature of $V_{GSR} = 130 \pm 10$ km s$^{-1}$ for members of the VOD/S297+63-20.5, which is extremely close to the VSS peak. The VSS and S297+63-20.5 are co-incident in space, but their velocity difference has not yet been reconciled. The distance measurements between S297+63-20.5 and the VSS are similar enough (to within $\sim1$ kpc) within probable distance uncertainties for \citet{Newberg;et-al_2007} and \citet{Prior;et-al_2009a} to infer they are part of the same structure. \citet{Newberg;et-al_2007} estimate a distance to S297+63-20.5 of $r_\odot = 18$ kpc from $g_0 = 20.5$ turnoff stars, but they concede the structure is likely dispersed along the line of sight as the Color-Magnitude Diagram (CMD) for this region does not demonstrate a tight sequence.  Certainly this region of sky, aptly coined the 'Field of Streams' by \citet{Belokurov;et-al_2006}, is complex territory.

Photometric studies are inadequate to fully untangle this region. Kinematics are essential to identify co-moving groups that are distinct from the general halo field. Chemical information is vital to accurately distinguish these substructures and understand their origins. However, very few studies have directly investigated metallicities for these stars. In this paper we present spectroscopic observations of K-giants in this region.  Substructure kinematics are used to probe the shape of dark matter in the Galaxy. We report both the velocities and metallicities of these giants in an effort to help untangle this accretion-dominated region.

Target selection methodology is outlined in the next section, which is followed with details regarding the observations. Techniques used to separate K-giants from dwarfs are discussed in \S\ref{sec:dwarf-giant-separation}, and our analysis procedure for kinematics (\S\ref{sec:kinematics}) and metallicities (\S\ref{sec:metallicities}) follows. A discussion of substructures is outlined in \S\ref{sec:discussion} and in \S\ref{sec:conclusions} we conclude with some final remarks and critical interpretations.

@@

When the presence of a stellar substructure is uncovered, K-giants provide excellent candidates for spectroscopic follow-up. They allow for precise radial velocities and chemical abundances. In order to specifically target K-giants, we have chosen candidates within the color selection box shown in Figure \ref{fig:cmd-target-selection}, taken from the SDSS DR7 catalogue. Field dwarfs are expected to contaminate the sample due to their similarity in colors. Although K-dwarfs are difficult to distinguish photometrically, we can spectroscopically separate these through the equivalent width (EW) of the  gravity-sensitive triplet Mg I lines at 5167.3, 5172.7, and 5183.6 \AA\ (see \S\ref{sec:dwarf-giant-separation}).

\begin{figure}[h]
	\includegraphics[width=\columnwidth]{./cmd_selection.eps}
	\caption{CMD of our observed regions from SDSS DR7, overlaid with the color selection criterion used to target K-giants. Appropriate Dartmouth isochrones \citep{Dotter;et-al_2008} are shown for Sgr debris at a distance of $r_\odot =40$ kpc \citep{Belokurov;et-al_2006}}
	\label{fig:cmd-target-selection}
\end{figure}

@@

Our targets were observed over two runs using AAOmega on the 3.9-m Anglo-Australian Telescope at Siding Springs Observatory in New South Wales, Australia. AAOmega is a double-beam, multi-object (392) fibre-fed spectrograph covering a two degree field of view. The targets were observed in normal visitor mode in April 2009. Throughout all observations, sufficient sky and guide fibres ($\sim30$ and 7-8 respectively) were allocated to ensure optimal sky subtraction and astrometry. In total 3,453 science targets were observed within 4 fields throughout the VOD/VSS region, as outlined in Table \ref{tab:observations}. Multiple configurations were observed for most fields to permit measurements of bright ($i < 16$) and faint ($i > 16$) stars, as well as repeat observations on a subset of stars.

@@

The beam was split into the red and blue arms using the 5700 \AA{} dichroic. The 580V grating in the blue arm yields spectra between 370-580 nm, with a resolution of $R = 1300$. In the red arm we used the 1000I grating ($R = 4400$) which spans the spectral range from 800-950 nm. This coverage includes the Ca II NIR Triplet (CaT), which is used for radial velocities and metallicities. To minimise scattered-light cross talk between fibres, science targets on each configuration were limited to 1.5 magnitudes in range. Globular clusters NGC 5024, 5053 and 5904 were observed as radial velocity and metallicity standards.

The data was reduced using the 2\textsc{DFDR}\footnote{http://www.aao.gov.au/AAO/2df/aaomega} pipeline. After being flat-fielded, the fibres were throughput calibrated and the sky spectrum was subtracted using the median of the dedicated sky fibres. Wavelength calibration was achieved from arc lamp exposures taken between each set of science fields. Multiple object frames were successively taken to assist with cosmic ray removal.

@@

	When discussing our data with respect to stellar streams and substructures within the halo, we are referring only to K-type giants.  Dwarfs that fall within our apparent magnitude limit are not sufficiently distant to probe halo substructures. Our resolution is adequate such that the Mg I triplet lines can be individually measured and used to discriminate dwarfs.

A grid of synthetic spectra has been generated to quantitively establish a giant/dwarf star separation criterion. The grid was generated using \citet{Castelli;Kurucz_2004} model atmospheres with MOOG\footnote{http://www.as.utexas.edu/$\sim$chris/moog.html} and the line list of \citet{Kurucz;Bell_1995}. Spectra was also generated using stellar parameters for the Sun and Arcturus. The strength of the Mg I lines were tuned to match both the Solar and Arcturus atlases of \citet{Hinkle;et-al_2003}. \citet{Girardi;et-al_2004} isochrones have been used to translate our de-reddened $g - i$ color range to effective temperature. Reddening is accounted for using the maps of \citet{Schlegel;Finkbeiner;Davis_1998}. The corresponding effective temperature region ranges from 3900 to 5200 K, and is stepped at 25 K intervals. We have assumed typical K-type surface gravities of $\log{g} = 2$ for giants and $\log{g} = 4.5$ for dwarfs. Metallicities of [Fe/H] $= -0.5$, $-1.5$, and $-2.5$ were considered for both surface gravities.

All synthetic spectra was mapped onto the same wavelength space as our observations. Intensity was convolved with a Gaussian kernel of 3.03 \AA{} to match our observed resolution with the 580V grating. Mg I line strengths for our observations and synthetic spectra is shown against $g - i$ in Figure \ref{fig:dwarf-giant-separation}. As expected, there is an overlap of Mg I strengths between metal-rich ``giants'' ($\log{g} = 2$) and metal-poor ``dwarfs'' ($\log{g} = 4.5$). However, we do not expect metal-poor dwarfs to be a principle contaminant due to their intrinsically low luminosities and paucity.

>>>>

\author{Andrew R. Casey, Stefan C. Keller and Gary Da Costa}
\affil{Research School of Astronomy \& Astrophysics,
	Australian National University, Mount Stromlo Observatory, via Cotter Rd, Weston, ACT 2611, Australia}

\begin{abstract}
We present observations obtained with the AAT's 2dF wide field spectrograph AAOmega of K-type stars located within a region of the sky which contains the Virgo Over-Density and the leading arm of the Sagittarius Stream.  On the basis of the resulting velocity histogram we isolate halo substructures in these overlapping regions including Sagittarius and previously discovered Virgo groups. Through comparisons with \textit{N}-body models of the Galaxy-Sagittarius interaction, we find a tri-axial dark matter halo is favoured and we exclude a prolate shape. This result is contradictory with other observations along the Sagittarius leading arm, which typically favour prolate models.  We have also uncovered K-giant members of Sagittarius that are notably more metal poor ($\langle[$Fe/H$]\rangle = -1.7\,\pm\,0.3$ dex) than previous studies. This suggests a significantly wider metallicity distribution exists in the Sagittarius Stream than formerly considered. We also present data on five carbon stars which were discovered in our sample.
\end{abstract}

\keywords{Galaxy: halo, structure --- Individual: Sagittarius, Virgo Stellar Stream --- Stars: carbon, K-giants}

@@

The proportion of substructure recently uncovered in the Galaxy has highlighted the crucial involvement accretion has played in the formation of the Milky Way \citep[see][and references therein]{Helmi_2008}. Properties of these stellar structures allow us to probe the formation mechanisms and history of the Galaxy. Recent studies \citep{Carollo;et-al_2007, Carollo;et-al_2010} have suggested multiple evolutionary paths are required for galaxy formation to reconcile observational evidence, although this is a subject of ongoing debate \citep{Schoenrich;et-al_2011}. Regardless, the dissipation-less merging paradigm is widely accepted and consistent with favoured Cold Dark Matter ($\Lambda$CDM) cosmology models. Through the examination of ongoing accretion events in the Milky Way and fossils from previous mergers we can trace the evolution of the outer most regions of the Galaxy \citep[e.g.,][]{Helmi;White_2001}.

@@

Unquestionably the most prominent ongoing accretion event within the Milky Way is that of the Sagittarius (Sgr) dwarf Spheroidal (dSph) galaxy. Originally discovered by \citet{Ibata;et-al_1994} as a co-moving group of K- and M-type giants, the tidal tails of Sgr circle our Galaxy. As such they have been extensively traced with red-clump stars \citep{Majewski;et-al_1999}, carbon stars \citep{Totten;Irwin_1998, Ibata;et-al_2001}, RR Lyrae stars \citep{Ivezic;et-al_2000, Vivas;et-al_2005, Keller;et-al_2008, Watkins;et-al_2009, Prior;et-al_2009b}, A-type stars \citep{Newberg;et-al_2003}, BHB stars \citep{Ruhland;et-al_2011} and K/M-giants \citep{Majewski;et-al_2003,  Yanny;et-al_2009, Keller;Yong;Da_Costa_2010}. Tracers originating from the host system can be unequivocally identified with spatial and kinematical information because they remain dynamically cold, and are identifiable as kinematic substructures long after they are stripped from their progenitor \citep[for example][]{Ibata;Lewis_1998, Helmi;White_1999}.

Stellar tracers within these tidal tails are kinematically sensitive to the galactic potential. This has led various groups to model the Sgr interaction with different dark matter profiles. \citet{Martinez-Delgado;et-al_2004} traced the Northern leading arm and found a near spherical or oblate ($q \approx 0.85$) dark matter halo best represented the observed debris, coinciding with the findings of \citet{Ibata;et-al_2001}. In contrast, \citet{Helmi_2004} found evidence in the Sgr leading debris that most favoured a prolate ($q = 1.25$) halo. \citet{Vivas;et-al_2005} found that either a prolate or spherical model of  \citet{Helmi_2004} would fit their RR Lyrae observations, rather than those of an oblate model. \citet{Johnston;et-al_2005} later pointed out that no prolate model can reproduce the orbital pole precession of the Sgr debris but an oblate potential could. \citet[hereafter LJM05]{Law;et-al_2005} performed simulations using data of the Sgr debris from the 2-Micron All Sky Survey (2MASS) catalogue and found that the kinematics of leading debris was best fit by prolate halos, whereas the trailing debris typically favoured oblate halos. \citet{Prior;et-al_2009a} reached a similar conclusion.

\citet{Belokurov;et-al_2006} found an apparent bifurcation within the Sgr debris which \citet{Fellhauer;et-al_2006} argued can only result from a dark halo having a near spherical shape. \citet{Law;et-al_2009} introduced a tri-axial model with a varying flattening profile $q$, which replicates the orbital precession seen and matches kinematic observations of the Sgr debris. However \citet{Law;Majewski_2010} (hereafter LM10) concede this may be a purely numerical solution as tri-axial halos are dynamically unstable. They emphasize that  more kinematic measurements in other regions of the Sgr stream are required.

After the Sgr stellar stream, the Virgo Over-Density (VOD) is arguably the next most significant substructure within our Galaxy. The first over-density in the vicinity of the VOD was observed as a group of RR Lyrae stars by the QUEST survey \citep{Vivas;et-al_2001}. The collaboration later named this the ``$12\fh4$ clump'' \citep{Zinn;et-al_2004}. The broad nature of the VOD was later uncovered from the SDSS catalogue as a diffuse over-density of main-sequence turnoff stars centered at $r_\odot \sim18$ kpc \citep[which ][dubbed as S297+63-20.5]{Newberg;et-al_2002}. The nomenclature on the substructure names within this region is varied, however in this paper when referring to the VOD we are discussing the spatial over-density of stars within the region, separate from any detected co-moving groups.

The difficulty arises in accurately distinguishing the VOD as there are multiple substructures along this line-of-sight. \citet{Duffau;et-al_2006} took observations of BHB and RR Lyrae within the ``$12\fh4$ clump'' and found a common velocity of $V_{GSR} = 99.8$ km s$^{-1}$ with $\sigma_{v} = 17.3$ km s$^{-1}$. It should be noted that the dispersion in kinematics measured by \citet{Duffau;et-al_2006} is essentially their velocity precision, so the substructure may possess a much smaller kinematic dispersion. This co-moving group was coined the Virgo Stellar Stream (VSS) to differentiate it from the broad spatial over-density. This distinction from the VOD was somewhat strengthened with new distance measurements which placed the VOD centroid at $r_\odot = 16$ kpc \citep{Juric;et-al_2008, Keller_2010}, and the VSS 3 kpc further away \citep{Duffau;et-al_2006}. Although, when considered in the light of systematic and observational uncertainties, this is of marginal significance. Additionally, \citet{Juric;et-al_2008} suggests the VOD may extend between $r_\odot = 6$ to 20 kpc, which further complicates the matter of distance separation.

The relationship between the VSS and the S297+63-20.5 over-density is still unclear. \citet{Newberg;et-al_2007} found a kinematic signature at $V_{GSR} = 130 \pm 10$ km s$^{-1}$ for members of the VOD/S297+63-20.5, which is extremely close to the VSS peak. The VSS and S297+63-20.5 are co-incident in space, but their velocity difference has not yet been reconciled. The distance measurements between S297+63-20.5 and the VSS are similar enough ($\sim1$ kpc) within probable distance uncertainties for \citet{Newberg;et-al_2007} and \citet{Prior;et-al_2009a} to infer they are part of the same structure. \citet{Newberg;et-al_2007} estimate a distance to S297+63-20.5 of $r_\odot = 18$ kpc from $g_0 = 20.5$ turnoff stars, but they concede the structure is likely dispersed along the line-of-sight as the Color-Magnitude Diagram (CMD) for this region does not demonstrate a tight sequence.  Certainly this region of sky, aptly coined the 'Field of Streams' by \citet{Belokurov;et-al_2006}, is complex territory.

Photometric studies are inadequate to fully untangle this region. Kinematics are essential to identify co-moving groups that are distinct from the general halo field. Chemical information is vital to accurately distinguish these substructures and understand their origins. However, very few studies have directly investigated metallicities for these stars. In this paper we present spectroscopic observations of K-giants in this region.  Sgr stream kinematics are used to probe the shape of the dark matter halo in the Galaxy. We report both the velocities and metallicities of these giants in an effort to help untangle this accretion-dominated region.

Target selection methodology is outlined in the next section which is followed with details regarding the observations. Techniques used to separate K-giants from dwarfs are discussed in \S\ref{sec:dwarf-giant-separation}, and our analysis procedure for kinematics (\S\ref{sec:kinematics}) and metallicities (\S\ref{sec:metallicities}) follows. A discussion of substructures is outlined in \S\ref{sec:discussion}, and in \S\ref{sec:carbon-stars} we report the carbon stars discovered in our sample. In \S\ref{sec:conclusions} we conclude with some final remarks and critical interpretations.

@@

When the presence of a stellar substructure is uncovered, K-giants provide excellent candidates for spectroscopic follow-up. They allow for precise radial velocities and chemical abundances. In order to specifically target K-giants we have chosen candidates within the color selection box shown in Figure \ref{fig:cmd-target-selection}, taken from the SDSS DR7 catalogue. This selection box favors a metal-poor population at Sgr stream-like distances ($\sim40$ kpc) but will also contain similarly metal-poor stars at VOD/VSS distances ($\sim20$ kpc) without being significantly biased. Field dwarfs are expected to contaminate the sample due to their similarity in colors. Although K-dwarfs are difficult to distinguish photometrically we can spectroscopically separate these through the equivalent width (EW) of the  gravity-sensitive Mg I triplet lines at 5167.3, 5172.7, and 5183.6 \AA\ (see \S\ref{sec:dwarf-giant-separation}).

\begin{figure}[h]
	\includegraphics[width=\columnwidth]{./cmd.eps}
	\caption{CMD of our observed regions from SDSS DR7, overlaid with the color selection criterion used to target K-giants. Appropriate Dartmouth isochrones \citep{Dotter;et-al_2008} are shown for Sgr debris at a distance of $r_\odot =40$ kpc \citep{Belokurov;et-al_2006}}
	\label{fig:cmd-target-selection}
\end{figure}

@@

Our targets were observed over two runs using AAOmega on the 3.9-m Anglo-Australian Telescope at Siding Springs Observatory in New South Wales, Australia. AAOmega is a double-beam, multi-object (392) fibre-fed spectrograph covering a two degree field of view. The targets were observed in normal visitor mode in April 2009. Throughout all observations, sufficient sky fibres ($\sim30$) were allocated to ensure optimal sky subtraction. In total 3,453 science targets were observed across 4 fields within the VOD/Sgr region, as outlined in Table \ref{tab:observations}. Multiple configurations were observed for most fields to permit measurements of bright ($i < 16$) and faint ($i > 16$) stars, as well as repeat observations on a subset of stars.

@@

The beam was split into the red and blue arms using the 5700 \AA{} dichroic. The 580V grating in the blue arm yields spectra between 370-580 nm, with a resolution of $R = 1300$. In the red arm we used the 1000I grating ($R = 4400$) which spans the spectral range from 800-950 nm. This coverage includes the Ca II NIR Triplet (CaT), which is used for radial velocities and metallicities. Science targets on each configuration were limited to 1.5 magnitudes in range to minimise scattered-light cross talk between fibres. Globular clusters NGC 5024, 5053 and 5904 were observed as radial velocity and metallicity standards.

The data was reduced using the 2\textsc{DFDR}\footnote{http://www.aao.gov.au/AAO/2df/aaomega} pipeline. After being flat-fielded, the fibres were throughput calibrated and the sky spectrum was subtracted using the median flux of the dedicated sky fibres. Wavelength calibration was achieved from arc lamp exposures taken between each set of science fields. Multiple object frames were combined to assist with cosmic ray removal.

@@

	When discussing our data with respect to stellar streams and substructures within the halo, we are referring only to K-type giants.  Dwarfs that fall within our apparent magnitude limit are not sufficiently distant to probe halo substructures. Our resolution is adequate such that the Mg I triplet lines can be individually measured and used to discriminate against dwarfs.

A grid of synthetic spectra has been generated to quantitively establish a suitable giant/dwarf separation criterion. The grid was generated using \citet{Castelli;Kurucz_2004} model atmospheres with MOOG\footnote{http://www.as.utexas.edu/$\sim$chris/moog.html} and the line list of \citet{Kurucz;Bell_1995}. Spectra were also generated using stellar parameters for the Sun and Arcturus. The strength of the Mg I lines were tuned to match both the Solar and Arcturus atlases of \citet{Hinkle;et-al_2003}. \citet{Girardi;et-al_2004} isochrones have been used to translate our de-reddened $g - i$ color range to effective temperature. Reddening is accounted for using the \citet{Schlafly;Finkbeiner_2011} corrected dust maps of \citet{Schlegel;Finkbeiner;Davis_1998} assuming a \citet{Fitzpatrick_1999} dust profile where $R_V = 3.1$. The corresponding effective temperature region ranges from 3900 to 5200 K, and is stepped at 25 K intervals. We have assumed typical K-type surface gravities of $\log{g} = 2$ for giants and $\log{g} = 4.5$ for dwarfs. Metallicities of [Fe/H] $= -0.5$, $-1.5$, and $-2.5$ were considered for both surface gravities.

All synthetic spectra was mapped onto the same wavelength intervals before the flux was convolved with a Gaussian kernel of 3.03 \AA{} to match our 580V observations. Mg I line strengths for our observations and synthetic spectra are shown against $g - i$ in Figure \ref{fig:dwarf-giant-separation}. As expected, there is an overlap of Mg I strengths between metal-rich ``giants'' ($\log{g} = 2$) and metal-poor ``dwarfs'' ($\log{g} = 4.5$). However we do not expect metal-poor dwarfs to be a principle contaminant due to their intrinsically low luminosities and comparative rarity.
